---
title: "Assignment 2"
author: "SABIHA MHATARNAIK"
date: "10/2/2021"
output: pdf_document
---
#Install lpSolveAPI package if not installed
```{r}
#install.packages("lpSolveAPI")
```

```{r}
#Now, load the library
library(lpSolveAPI)

#create an lp object named 'lprec' with 0 constraints and 9 decision variables
lprec <- make.lp(0,9)

#Now create the objective function. The default is a minimization problem.
set.objfn(lprec, c(420,420,420,
                   360,360,360,
                   300,300,300))

# As the default is a minimization problem, we change the direction to set maximization
lp.control(lprec,sense='max')

#Add the 12 constraints based on the plant's number and products made on those plants.
add.constraint(lprec ,c(1,0,0,1,0,0,1,0,0), "<=", 750)
add.constraint(lprec ,c(0,1,0,0,1,0,0,1,0), "<=", 900)
add.constraint(lprec ,c(0,0,1,0,0,1,0,0,1), "<=", 450)

add.constraint(lprec ,c(20,0,0,15,0,0,12,0,0), "<=", 13000)
add.constraint(lprec ,c(0,20,0,0,15,0,0,12,0), "<=", 12000)
add.constraint(lprec ,c(0,0,20,0,0,15,0,0,12), "<=", 5000)

add.constraint(lprec ,c(1,1,1,0,0,0,0,0,0), "<=", 900)
add.constraint(lprec ,c(0,0,0,1,1,1,0,0,0), "<=", 1200)
add.constraint(lprec ,c(0,0,0,0,0,0,1,1,1), "<=", 750)

add.constraint(lprec ,c(900,-750,0,900,-750,0,900,-750,0), "=", 0)
add.constraint(lprec ,c(0,450,-900,0,450,-900,0,450,-900), "=", 0)
add.constraint(lprec,c(450,0,-750,450,0,-750,450,0,-750),"=",0)
```
Set bounds for variables.

Remember that all variables had to be non-negative. We don't need to do it 
here,as this is the default , we can set bounds explicitly.

```{r}
#Set bounds for variables.
set.bounds(lprec ,lower =c(0,0,0,0,0,0,0,0,0),
           columns= c (1:9))  #Not really needed

# To identify  the variables and constraints, we can set variable names and constraint names.
RowNames <-c("P1ProductionCapacity","P2ProductionCapacity","P3ProductionCapacity",
             "P1StorageSpace","P2StorageSpace","P3StorageSpace",
             "SalesForecastLarge","SalesForecastMedium","SalesForecastSmall",
             "PercentCapacityP1andP2","PercentCapacityP2andP3","PercentCapacityP1andP3")

ColNames <- c("Plant1Large","Plant2Large","Plant3Large",
              "Plant1Medium","Plant2Medium","Plant3Medium",
              "Plant1Small","Plant2Small","Plant3Small")

dimnames(lprec)<- list (RowNames,ColNames)

#Now view the Model
lprec

```

```{r}
# The model can also be saved to a file
write.lp(lprec, filename = "weigelt.lp", type = "lp")
```

Now we can solve the above LP Problem
```{r}
solve(lprec)
```
The output above indicated that the answer is 0, means there was a successful solution.
We now output the value of the objective function, and the variables.

```{r}
get.objective(lprec)
get.variables(lprec)

```
From the above solution, we can infer the following :
Plant 1 : 516.67 of Large Products and 177.78 of Medium Products.
Plant 2 : 666.67 of Medium Products and 166.67 of Small products.
Plant 3 : 416.67 of Small Products

```{r}
get.constraints(lprec)
```

We now read the lp formulation using an lp file.
I am using the same R file which I have saved.

```{r}

a <- read.lp ("weigelt.lp")  # create an lp object a
a                            #display a
```

Solve the lp model
```{r}
solve(a)                     #get objective value
get.objective(a)             #get values of decision variables
get.constraints(a)           #get constraint values
```
