---
title: "Assignment 1"
author: "SABIHA MHATARNAIK"
date: "9/4/2021"
output:
  pdf_document: default
  word_document: default
---

##Set the working directory
```{r}
setwd( "~/Desktop/Fall 2021/QMM/Assignment 1")
```

#Reading a file for data
```{r}
Dataset_Media <- read.csv("Dataset_Media.csv")
```
#head function gives the first 6 rows
```{r}
head(Dataset_Media)
```

#tail function gives the last 6 rows 
```{r}
tail(Dataset_Media)
```

#str function will structure the data 
```{r}
str(Dataset_Media)
```

#summary function will give the summary of the data 
```{r}
summary(Dataset_Media)
```

```{r}
plot(Dataset_Media)
```
#X-Y plot for two quantitative variables(Scatterplot)
```{r}
plot(Dataset_Media$Audience_Size,Dataset_Media$Preference)
```
#Add some Options
```{r}
plot(Dataset_Media$Audience_Size,Dataset_Media$Preference,pch=20,col="#cc0000",main="Example of Media",xlab="Size of the Audience",ylab="Preference")
```
#Plotting a quantitative variable Audience_Size
```{r}
plot(Dataset_Media$Audience_Size)
```
#Lineplot,Histogram,Boxplot
```{r}
plot(Dataset_Media$Audience_Size,type="l",main="Example of Media",xlab="Size of the Audience",ylab="Preference")
hist(Dataset_Media$Audience_Size)
boxplot(Dataset_Media$Audience_Size)
```
#Working with functions
```{r}
min(Dataset_Media$Audience_Size)
max(Dataset_Media$Audience_Size)
range(Dataset_Media$Audience_Size)
mean(Dataset_Media$Audience_Size)
median(Dataset_Media$Audience_Size)
```