---
title: "Assignment 4"
author: "SABIHA MHATARNAIK"
date: "10/23/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Q.1)
***Solving this transportation problem utilizing R***


**Adding the LP Objective function and constraint**

We will be using "lpSolveAPI" package in R
```{r}
#Import the lpSolveAPI package
library(lpSolveAPI)
```

```{r}
#Creating an lp object named lptrans with 0 constraints and 6 decision variables.
lptrans <- make.lp(0,6)

#Set the objective function.Since,minimization problem is the default so we don't 
#have to change the direction.
set.objfn(lptrans, c(622,614,630,
                     641,645,649))

#Add the 9 constraints based on the number of units made on each plant and 
#shipped to different warehouses.

add.constraint(lptrans, c(1,1,1,0,0,0), "<=", 100)
add.constraint(lptrans, c(0,0,0,1,1,1), "<=", 120)

add.constraint(lptrans, c(1,0,0,1,0,0), "=", 80)
add.constraint(lptrans, c(0,1,0,0,1,0), "=", 60)
add.constraint(lptrans, c(0,0,1,0,0,1), "=", 70)
```

Setting the bounds for the variables.
Also, it is important to note that all variables have to be non-negative.
There is no need to do this as this is by default,but we can set the bounds explicitly.

```{r}
#Set bounds for variables.
set.bounds(lptrans ,lower =c(0,0,0,0,0,0),
           columns= c(1:6)) # not really needed

#To identify the variables and constraints, we can
# set variable names and name the constraints
RowNames <- c ("PrdtnCapPlantA" ,"PrdctnCapPlantB",
               "DemandWh1","DemandWh2","DemandWh3")

ColNames <- c("PlantAWh1","PlantAWh2","PlantAWh3",
              "PlantBWh1","PlantBWh2","PlantBWh3")

dimnames(lptrans) <- list(RowNames,ColNames)

#Now print out the model 
lptrans
```
```{r}
#Saving to a file
write.lp(lptrans,filename="AED.lp",type="lp")
```
**Now solve the lp problem**
```{r}
solve(lptrans)
```
The output above indicate that the result is 0, means that there was a successful solution
We now output the value of the objective function, and the variables.

```{r}
#It will give us the minimum value for the optimal solution.
get.objective(lptrans)
```
Hence $132790 is the minimum combined cost of production and shipping for the optimal solution.

```{r}
#It will give us the optimal number of units produced on each plant and shipped to each warehouse.
get.variables(lptrans)
```
In order to get the minimum combined cost of production and shipping each plant should produce and ship the following amounts:
Plant A should produce 100 units. the company should ship 60 units of plant A to warehouse 2 and 40 units to warehouse 3.
Plant B should produce 110 units .The company should ship 80 units of plant B to warehouse 1 and 30 units
to warehouse 3

```{r}
#It will give us the optimal constraints
get.constraints(lptrans)
```

This output is important that shows the correct values in order to satisfy the equality assumption.The
equality assumption in a transportation problem says that the given problem will have a feasible solution iff total production= total demand.
Hence, the constraints on plant B must be reduced from 120 to 110 to have a feasible solution.

```{r}
get.sensitivity.obj(lptrans) #get Reduced Costs
```

```{r}
get.sensitivity.rhs(lptrans) #get Shadow Price
```

**Dummy Variable Approach**

```{r}
#Creating an lp object named lptrans with 0 constraints and 8 decision variables.
lptrans_dummy<- make.lp(0,8)

#Set the objective function.Since,minimization problem is the default so we don't 
#have to change the direction.
set.objfn(lptrans_dummy, c(622,614,630,0,
                     641,645,649,0))
```

Add the 8 constraints based on the number of units made on each plant and 
shipped to different warehouses and the dummy variable that is needed which will be a dummy warehouse 4
to satisfy the equality assumption i.e total production = total demand.
```{r}
add.constraint(lptrans_dummy, c(1,1,1,1,0,0,0,0), "=", 100)
add.constraint(lptrans_dummy, c(0,0,0,0,1,1,1,1), "=", 120)

add.constraint(lptrans_dummy, c(1,0,0,0,1,0,0,0), "=", 80)
add.constraint(lptrans_dummy, c(0,1,0,0,0,1,0,0), "=", 60)
add.constraint(lptrans_dummy, c(0,0,1,0,0,0,1,0), "=", 70)
add.constraint(lptrans_dummy, c(0,0,0,1,0,0,0,1), "=", 10) #Dummy Variable

#Set bounds for variables.
set.bounds(lptrans_dummy ,lower =c(0,0,0,0,0,0,0,0),
           columns= c(1:8)) # not really needed

#To identify the variables and constraints, we can
# set variable names and name the constraints
RowNames_dummy <- c ("PrdtnCapPltA" ,"PrdctnCapPltB",
               "DemandWh1","DemandWh2","DemandWh3","DemandWh4")

ColNames_dummy <- c("PltAWh1","PltAWh2","PltAWh3","PltAWh4",
              "PltBWh1","PltBWh2","PltBWh3","PltBWh4")

dimnames(lptrans_dummy) <- list(RowNames_dummy,ColNames_dummy)

#Now print out the model 
lptrans_dummy
```
```{r}
#Solve the transportation problem 
solve(lptrans_dummy)
```
The output above indicate that the result is 0, means that there was a successful solution.

```{r}
#It will give us the minimum value for the optimal solution.
get.objective(lptrans_dummy)
```

Same optimal solution as that of the previous approach

```{r}
#It will give us the optimal number of units produced on each plant and shipped to each warehouse.
get.variables(lptrans_dummy)
```
Here, the additional 10 units are logically sent to warehouse 4,which is needed to satisfy th equality assumption.Warehouse 4 is not a physical location, but it means 10 units must be shipped to any warehouse to get a feasible solution.

```{r}
#It will give us the optimal constraints
get.constraints(lptrans_dummy)
```
Here, all constraints are satisfied.

```{r}
get.sensitivity.obj(lptrans_dummy) # Reduced Cost
```
```{r}
#It will give us the optimal constraints
get.sensitivity.rhs(lptrans_dummy) # Shadow Price
```
***Also, we can read the lp formulation using an lp file and solve it.***
```{r}
x <- read.lp("AED.lp")
x
```

```{r}
solve(x)
get.objective(x) #get the objective value
get.variables(x) #get the values f decision variables
get.constraints(x) #get constraints 
```

Q.2)
***Oil Distribution***

1) 

***Solving this transportation problem utilizing R***

```{r}
#Creating an lp object named lptrans with 0 constraints and 27 decision variables.
lptransship<-make.lp(0,27)
lp.control(lptransship,sense='min')
#Set the objective function.
set.objfn(lptransship,c(1.52,1.60,1.40,1.70,1.63,1.55,1.45,1.57,1.30,5.15,5.12,5.32,5.69,5.47,6.16,6.13,6.05,6.25,5.63,6.12,6.17,5.80,5.71,5.87,0,0,0))
#Adding the constraints
add.constraint(lptransship,c(1,1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),"=",93)
add.constraint(lptransship,c(0,0,0,1,1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),"=",88)
add.constraint(lptransship,c(0,0,0,0,0,0,1,1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),"=",95)
add.constraint(lptransship,c(0,0,0,0,0,0,0,0,0,1,1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),"=",30)
add.constraint(lptransship,c(0,0,0,0,0,0,0,0,0,0,0,0,1,1,1,0,0,0,0,0,0,0,0,0,0,0,0),"=",57)
add.constraint(lptransship,c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,1,1,0,0,0,0,0,0,0,0,0),"=",48)
add.constraint(lptransship,c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,1,1,0,0,0,0,0,0),"=",91)
add.constraint(lptransship,c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,1,1,0,0,0),"=",48)
add.constraint(lptransship,c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,1,1),"=",2)
add.constraint(lptransship,c(1,0,0,1,0,0,1,0,0,-1,0,0,-1,0,0,-1,0,0,-1,0,0,-1,0,0,-1,0,0),"=",0)
add.constraint(lptransship,c(0,1,0,0,1,0,0,1,0,0,-1,0,0,-1,0,0,-1,0,0,-1,0,0,-1,0,0,-1,0),"=",0)
add.constraint(lptransship,c(0,0,1,0,0,1,0,0,1,0,0,-1,0,0,-1,0,0,-1,0,0,-1,0,0,-1,0,0,-1),"=",0)

solve(lptransship)
get.objective(lptransship)
get.constraints(lptransship)
get.variables(lptransship)
```


