---
title: "Assignment 3"
author: "SABIHA MHATARNAIK"
date: "10/11/2021"
output:
  pdf_document: default
  html_document: default
---


Q1. Solve the problem using lpSolve or any other equivalent library in R.
#Install lpSolveAPI package if not installed
```{r}
#install.packages("lpSolveAPI")
```

```{r}
#Now, load the library
library(lpSolveAPI)

#create an lp object named 'lprec' with 0 constraints and 9 decision variables
lprec <- make.lp(0,9)

#Now create the objective function. The default is a minimization problem.
set.objfn(lprec, c(420,420,420,
                   360,360,360,
                   300,300,300))

# As the default is a minimization problem, we change the direction to set maximization
lp.control(lprec,sense='max')

#Add the 12 constraints based on the plant's number and products made on those plants.
add.constraint(lprec ,c(1,0,0,1,0,0,1,0,0), "<=", 750)
add.constraint(lprec ,c(0,1,0,0,1,0,0,1,0), "<=", 900)
add.constraint(lprec ,c(0,0,1,0,0,1,0,0,1), "<=", 450)

add.constraint(lprec ,c(20,0,0,15,0,0,12,0,0), "<=", 13000)
add.constraint(lprec ,c(0,20,0,0,15,0,0,12,0), "<=", 12000)
add.constraint(lprec ,c(0,0,20,0,0,15,0,0,12), "<=", 5000)

add.constraint(lprec ,c(1,1,1,0,0,0,0,0,0), "<=", 900)
add.constraint(lprec ,c(0,0,0,1,1,1,0,0,0), "<=", 1200)
add.constraint(lprec ,c(0,0,0,0,0,0,1,1,1), "<=", 750)

add.constraint(lprec ,c(900,-750,0,900,-750,0,900,-750,0), "=", 0)
add.constraint(lprec ,c(0,450,-900,0,450,-900,0,450,-900), "=", 0)
add.constraint(lprec,c(450,0,-750,450,0,-750,450,0,-750),"=",0)
```
Set bounds for variables.

Remember that all variables had to be non-negative. We don't need to do it 
here,as this is the default , we can set bounds explicitly.

```{r}
#Set bounds for variables.
set.bounds(lprec ,lower =c(0,0,0,0,0,0,0,0,0),
           columns= c (1:9))  #Not really needed

# To identify  the variables and constraints, we can set variable names and constraint names.
RowNames <-c("P1ProductionCapacity","P2ProductionCapacity","P3ProductionCapacity",
             "P1StorageSpace","P2StorageSpace","P3StorageSpace",
             "SalesForecastLarge","SalesForecastMedium","SalesForecastSmall",
             "PercentCapacityP1andP2","PercentCapacityP2andP3","PercentCapacityP1andP3")

ColNames <- c("Plant1Large","Plant2Large","Plant3Large",
              "Plant1Medium","Plant2Medium","Plant3Medium",
              "Plant1Small","Plant2Small","Plant3Small")

dimnames(lprec)<- list (RowNames,ColNames)

#Now view the Model
lprec

```

```{r}
# The model can also be saved to a file
write.lp(lprec, filename = "weigelt.lp", type = "lp")
```

Now we can solve the above LP Problem
```{r}
solve(lprec)
```
The output above indicated that the answer is 0, means there was a successful solution.
We now output the value of the objective function, and the variables.

```{r}
get.objective(lprec)
get.variables(lprec)

```
From the above solution, we can infer the following :
Plant 1 : 516.67 of Large Products and 177.78 of Medium Products.
Plant 2 : 666.67 of Medium Products and 166.67 of Small products.
Plant 3 : 416.67 of Small Products

```{r}
get.constraints(lprec)
```

We now read the lp formulation using an lp file.
I am using the same R file which I have saved.

```{r}

a <- read.lp ("weigelt.lp")  # create an lp object a
a                            #display a
```

Solve the lp model
```{r}
solve(a)                     #get objective value
get.objective(a)             #get values of decision variables
get.constraints(a)           #get constraint values
get.variables(lprec)
```

Q2. Identify the shadow prices,dual solution and reduced costs.
```{r}
get.sensitivity.rhs(lprec) # get Shadow Prices
```
Note that the shadow prices are expressed here under $duals.
We also have valid ranges for shadow price calculations. Those are given under $dualsfrom and $dualstill.

```{r}
get.sensitivity.obj(lprec) # get Reduced Costs
```
The reduced costs are expressed here until $objfrom and $objtill.

Dual solution
```{r}
get.dual.solution(lprec )   #get dual Solution
```

Q3. Further,identify the sensitivity of the above prices and costs.That is,specify the range of shadow prices and reduced cost within which the optimal solution will not change

```{r}
cbind( get.sensitivity.rhs(lprec)$duals[1:21],lowerRange=get.sensitivity.rhs(lprec)$dualsfrom[1:21],upperRange=get.sensitivity.rhs(lprec)$dualstill[1:21])
```
The shadow prices are expressed here comprising of $dualsfrom and $dualstill.
The above is the range specified for the shadow prices within which the optimal solution will not change.

```{r}
cbind(get.sensitivity.obj(lprec)$duals[1:9],lowerRange=get.sensitivity.obj(lprec)$objfrom[1:9],upperRange=get.sensitivity.obj(lprec)$objtill[1:9])
```
The reduced costs are expressed here until $objfrom and $objtill.
The above is the range specified for the reduced cost within which the optimal solution will not change.


```{r}
get.sensitivity.rhs(lprec)$duals
```

Q4. Formulation of the dual of the above problem 
```{r}
lpDual <- make.lp(0,12)
```

```{r}
set.objfn(lpDual, c(750,900,450,
                   13000,12000,5000,
                   900,1200,750,
                   0,0,0))
```

```{r}
lp.control(lpDual,sense='min',simplextype="dual")
add.constraint(lpDual ,c(1,0,0,20,0,0,1,0,0,900,0,450), ">=", 420)
add.constraint(lpDual ,c(0,1,0,0,20,0,1,0,0,-750,450,0), ">=", 420)
add.constraint(lpDual ,c(0,0,1,0,0,20,1,0,0,0,-900,-750), ">=", 420)

add.constraint(lpDual ,c(1,0,0,15,0,0,0,1,0,900,0,450), ">=", 360)
add.constraint(lpDual ,c(0,1,0,0,15,0,0,1,0,-750,450,0), ">=", 360)
add.constraint(lpDual ,c(0,0,1,0,0,15,0,1,0,0,-900,-750), ">=", 360)

add.constraint(lpDual ,c(1,0,0,12,0,0,0,0,1,900,0,450), ">=", 300)
add.constraint(lpDual ,c(0,1,0,0,12,0,0,0,1,-750,450,0), ">=", 300)
add.constraint(lpDual ,c(0,0,1,0,0,12,0,0,1,0,-900,-750), ">=", 300)
```

We now read the lp formulation using an lp file
```{r}
lpDual <- read.lp("dual.lp") #Create an lp object lpDual
lpDual                     #Display lpDual
```

Solve the lp model
```{r}
solve(lpDual)                     #get objective value
get.objective(lpDual)             #get values of decision variables
get.variables(lpDual)
get.constraints(lpDual)           #get constraint values
get.sensitivity.rhs(lpDual)       #get shadow price
get.sensitivity.obj(lpDual)       #get reduced cost
get.dual.solution(lpDual)         #get dual solution

```

Specifying the range

```{r}
cbind( get.sensitivity.rhs(lpDual)$duals[1:21],lowerRange=get.sensitivity.rhs(lpDual)$dualsfrom[1:21],upperRange=get.sensitivity.rhs(lpDual)$dualstill[1:21])
```
The shadow prices are expressed here comprising of $dualsfrom and $dualstill.
The above is the range specified for the shadow prices within which the optimal solution will not change.

```{r}
cbind(get.sensitivity.obj(lpDual)$duals[1:12],lowerRange=get.sensitivity.obj(lpDual)$objfrom[1:12],upperRange=get.sensitivity.obj(lpDual)$objtill[1:12])
```
The reduced costs are expressed here until $objfrom and $objtill.
The above is the range specified for the reduced cost within which the optimal solution will not change.

The Formulation of dual yields results that agree with primal solution.

